Preliminary Insights into Indigenous Subsistence and Colonial Encounters: A Zooarchaeological Perspective from Kiwulan in Northeastern Taiwan

### 2025-06-04 analysis/paper/paper.qmd
revised based on comments

@tbl-ubiquity shows deer was the most widely used animal from CL6 to CL2, with the highest ubiquity in CL4 and CL5 (`r CL4_ubi_deer`), and the lowest in CL2. The second and third most commonly used animals are suid and muntjac. Although both reached their highest ubiquity in CL4 (`r CL4_ubi_boar` for suid, , and `r CL4_ubi_munt` for muntjac), their presence patterns differ slightly. Suid was utilized as early as CL6, declined in CL3, and absence in CL2, while muntjac showed a cultural lag, appearing one phase later than suid. Bovine had low ubiquity in CL5 and high ubiquity in CL1. Small animals (birds, fish, and turtle) had much lower ubiquity than large animals, with a slight increase in ubiquity in CL3. The result of ubiquity corresponds with the frequency data (NISP), suggesting that deer was the most common resource at the site, with an increased use of suid and muntjac in CL4.

>>>>

@tbl-ubiquity shows deer was the most widely used animal from CL6 to CL2, with the highest ubiquity in CL4 and CL5 (`r CL4_ubi_deer`), and the lowest in CL2. The second and third most commonly used animals are suid and muntjac. Although both reached their highest ubiquity in CL4 (`r CL4_ubi_boar` for suid, , and `r CL4_ubi_munt` for muntjac), their presence patterns differ slightly. Suid was utilized as early as CL6, declined in CL3, and absence in CL2, while muntjac showed a cultural lag, appearing one phase later than suid. Bovine had low ubiquity in CL5 and high ubiquity in CL1. Small animals (birds, fish, and turtle) had much lower ubiquity than large animals, with a slight increase in ubiquity in CL3. The results of the ubiquity analysis generally correspond with the frequency data (@fig-NISP-proportion: D), suggesting that deer was the most common resource at the site, with an increased use of suid in CL4. Although muntjac had relatively higher ubiquity in CL4 compared to other layers, their %NISP did not increase accordingly but instead decreased slightly, implying they may have been more widely used but only in small quantities.
